---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows
#-------------------------------------------------------------------

With SQL, you can filter columns with SELECT and rows with WHERE. Let's try an example with the Lahman baseball database. First, we need to load the appropriate packages. 

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```

Now, suppose we would like to see the homerun totals for the 1927 Yankees. We could write the following:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```


##### Continuing to practice writing queries and narrowing searches through multiple in-class examples-
#### Example One
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

#### Example Two 

```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting 
WHERE HR>=40 and SO<60"
sqldf(query)
```

#### Example Three

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE yearID>1969 and yearID<1980 and HR>30 and teamID='PHI'"
sqldf(query)
```

## Introducing 'Order By'
The command ORDER BY in a query can organize the information in the rows and columns by whichever numerical statistic you chose. 
It will default to listing the smallest number to largest, to reverse that type DESC at the end of the query.
In the example below we chose to list homeruns in decsending order- meaning the first on the list is the player with the most homeruns

#---------------------------------------------
#### Example One

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

#### Example Two
```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO,AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

##### Today in class, we reviewed previsouly taught queries- here is one example from Sept. 12

```{r}
query<-"SELECT playerID,HR FROM Batting
 WHERE playerID='ruthba01'"

sqldf(query)
```

## Aggregation
Aggregation is grouping. It is the narrowing of queries previously exercised. It is done with the command "GROUP BY" when using this you can group together any particular statiscal row, in the first example below we grouped by playerID, looking for Babe Ruth's summed up homeruns. The sum(_) command was also introduced simply adding stats together to organize the column,

#---------------------------------------------------

#### Example One

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf(query)
```

#### Example Two 
In this one it was discovered that after using the command GROUP BY if you are looking for the WHERE specification you use HAVING instead

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

#### Example Three
This one used avg() instead of sum- same idea

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting 
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf(query)
```

## Joins
The INNER JOIN command will join multiple tables together. Today we wanted to join columns with player information to the already used Batting table. Making the table produced more legible as in example one the players first and last name were presented rather than the playerID,

#---------------------------------------------------

#### Example One
Since the phrase playerID is present in both tables we needed to clarify where we wanted the playerID in the last line to come from
```{r}
query<-"SELECT namefirst,namelast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

#### Example Two

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

#### Example Three
In this one we had to clarify which table we wanted the teamID and yearID from. 
```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams 
ON Batting.teamID=teams.teamID AND Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)
```

#### Example Four
This will be the last thing for this website
```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
GROUP BY Salaries.playerID
HAVING nameFirst='Bob' AND avg(salary)>1000000"

sqldf(query)
```

